---
title: "Supplementary Material"
author: ""
date: "`r Sys.Date()`"
output:
  pdf_document:
    fig_caption: yes
  html_document:
    fig_caption: yes
---


All prices were converted in EUR before the analysis.

The most comprehensive longitudinal per publisher data sets are available from 
Canada (2014-2016), Chile (2008-2015), Finland (2010-2016), France (2009-2015), Netherlands (2011-2015), and UK (2010-2016). Data for Argentina (2008-2016) is also available but we could not verify the source. In most countries, the data is available only for some of the largest publishers and/or institutions.


```{r setup, echo=FALSE, message=FALSE, warning=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(devtools)
library(gridExtra)
library(phyloseq)
library(dplyr)
library(reshape2)
library(knitr)
library(plotROC)
library(pROC)
library(ggROC)
library(tidyr)
library(microbiome)
library(ggplot2)
#opts_chunk$set(fig.width=4, fig.height=3, par=TRUE, out.width='2in', fig.pos='H')
knitr::opts_chunk$set(fig.path = "figure_manuscript/", dev="CairoPNG")

# http://derekogle.com/fishR/2015-09-17-Figure-Table-Captions-in-Markdown
library(captioner)
tbls <- captioner(prefix="Table")
figs <- captioner(prefix="Figure")
# Check `r figs("ctrlsplit", display="num")`.

theme_set(theme_bw(20))
library(xtable)
library(ggplot2)
library(dplyr)
library(tidyr)
theme_set(theme_bw(20))

# Pick publishers that have time series (T>=3) in multiple (N>=3) countries
selected.publishers <- as.character(unlist((d %>%
  group_by(Country, Publisher) %>%
  tally() %>%
  filter(n>=3) %>% # Time points
  select(Country, Publisher) %>%
  group_by(Publisher) %>%
  tally() %>%
  filter(n>=3) %>% # Countries
  arrange(desc(n)))$Publisher))
```



```{r costs3, cache=FALSE, echo=FALSE, message=FALSE, warning=FALSE, fig.width=20, fig.height=10}
d2 <- d
dsub <- d2 %>% select("Country", "Year", "Publisher", "Cost") %>%
              filter(Publisher %in% selected.publishers) %>%
	      group_by(Country, Year, Publisher) %>%
	      summarize(Cost = sum(Cost, na.rm = TRUE))
# Mark zero costs as NA
dsub$Cost[dsub$Cost == 0] <- NA

# Check the largest publishers from here
my.pubs <- c("Elsevier", "Springer", "Wiley / Blackwell")
dsub <- dsub %>% filter(Publisher %in% my.pubs)

# Calculate cost increase (percent) per country/publisher and year
inc <- NULL
for (country in unique(dsub$Country)) {
  for (publisher in unique(dsub$Publisher)) {
    for (t in unique(subset(dsub, Country == country & Publisher == publisher)$Year)[-1]) {
      t1 <- t-1
      t2 <- t
      ind1 <- which(dsub$Country == country & dsub$Publisher == publisher & dsub$Year == t1)
      ind2 <- which(dsub$Country == country & dsub$Publisher == publisher & dsub$Year == t2)

      inc <- rbind(inc, c(ind2, dsub[ind2,"Cost"]/dsub[ind1,"Cost"] - 1))
    }
  }
}
inc <- cbind(unlist(inc[,1]), unlist(inc[,2]))
colnames(inc) <- c("index", "increase")
dsub$Increase <- rep(NA, nrow(dsub))
dsub$Increase[inc[, "index"]] <- inc[, "increase"]
dsub$Increase[is.infinite(dsub$Increase)] <- NA

library(scales)
colors <- c("black", "darkgray", "lightgray")
names(colors) <- my.pubs
pics <- list()
for (country in unique(dsub$Country)) {
  dd <- filter(dsub, Country == country)
  dd$Country <- droplevels(dd$Country) # factor(dd$Country, levels = levels(dsub$Country))
  dd$Publisher <- factor(dd$Publisher, levels = levels(dsub$Publisher))
  dd$Year <- as.factor(dd$Year)

  p <- ggplot(dd, aes(x = Publisher, y = Increase, fill = Year)) +
    geom_bar(stat = "identity", position = "dodge", color = "black", aes(group = Year)) + 
    labs(title = country, y = "Increase (%)") +
    scale_fill_grey()    
  pics[[country]] <- p
}
grid.arrange(pics[[1]], pics[[2]], pics[[3]], pics[[4]], pics[[5]], nrow = 2)
```

`r figs("costs3", "Annual cost increase for Elsevier, Wiley/Blackwell and Springer. In some data sets, the number of included institutions varies notably by year (for instance in Finland 2016 there is notable increase of listed publishers and organizations in the data set 57 vs. 71 organizations 2015-2016; and 200 vs. 258 publishers). However, the observations remain qualitatively similar also when the analysis is limited only to those publishers that have data available for the complete time span.")`.



```{r ranking1, cache=FALSE, echo=FALSE, message=FALSE, warning=FALSE, fig.width=14, fig.height=5, fig.show = "hold", out.width = "430px"}
# Ranking variables
# vars <- c("world_rank", "teaching", "international", "research", "citations", "income", "total_score", "num_students", "student_staff_ratio", "international_students")
vars <- c("world_rank", "research", "citations", "num_students")

# Merge cost and ranking data
drank <- inner_join(d, unirank, by = c("Year", "Country", "Organization"))

#p <- ggplot(subset(drank, Publisher %in% my.pubs), aes(x = world_rank, y = Cost, color = Country)) + geom_point() + labs(y = "Cost (EUR)") + scale_y_log10() + facet_grid(. ~ Publisher)

df <- subset(drank, Publisher %in% my.pubs);
res <- NULL
for (varname in vars) {

  df$ranking <- as.numeric(gsub(",", ".", as.character(df[[varname]])))
  r <- cor.test(log10(df$Cost), df$ranking, method = "spearman", use = "pairwise.complete.obs")
  lmres <- lm(log10(1 + Cost) ~ ranking, data = df)

  # Constant for Bonferroni correction
  bonferroni <- length(unique(df$Country)) * length(vars) * length(unique(df$Publisher))


  for (country in unique(df$Country)) {
    for (publisher in unique(df$Publisher)) {  
  
      dfsub <- subset(df, Country == country & Publisher == publisher)
      lmres <- lm(log10(1 + Cost) ~ ranking, data = dfsub)
      vec <- c()
      vec[["Variable"]] <- varname
      vec[["Country"]] <- country
      vec[["Publisher"]] <- publisher
      vec[["slope"]] <- coef(lmres)[["ranking"]]
      vec[["padj"]] <- anova(lmres)["ranking", "Pr(>F)"] * bonferroni
      res <- rbind(res, vec)
    }
  }

  p <- ggplot(df, aes(x = ranking, y = Cost, color = Country)) + geom_point() + labs(y = "Cost (EUR)", x = varname, title = varname) + scale_y_log10() + facet_grid(. ~ Publisher) + geom_smooth(method = "lm")
  print(p)
}
colnames(res) <- names(vec)
rownames(res) <- NULL
res <- as.data.frame(res)
res$slope <- as.numeric(as.character(res$slope))
res$padj <- as.numeric(as.character(res$padj))
```

`r figs("ranking1", "University ranking (Times ranking) variables versus subscription costs with selected publishers.")`


```{r ranking1tab, cache=FALSE, echo=FALSE, message=FALSE, warning=FALSE}
kable(res, digits = 6)
```

`r tbls("ranking1tab", "University ranking (Times ranking) variables versus subscription costs with selected publishers. Linear model slope and Bonferroni-adjusted p-value.")`

```{r rankingeter, cache=FALSE, echo=FALSE, message=FALSE, warning=FALSE, fig.width=15, fig.height=6, fig.show = "hold", out.width = "290px"}
# Ranking variables
vars <- as.character(eter.fields$code)

# Merge cost and ranking data
drank <- inner_join(d, eter, by = c("Year", "Organization"))

df <- subset(drank, Publisher %in% my.pubs);
for (varname in vars) {
  df$ranking <- df[[varname]];
  name <- as.character(subset(eter.fields, code == varname)$name)
  code <- varname
  p <- ggplot(df, aes(x = ranking, y = Cost, color = Country)) + geom_point() + labs(y = "Cost (EUR)", x = varname, title = name, subtitle = code) + scale_y_log10() + facet_grid(. ~ Publisher)
  print(p)
}
```

`r figs("rankingeter", "University scoring (ETER) variables versus subscription costs with selected publishers.")`






```{r costs1, cache=FALSE, echo=FALSE, message=FALSE, warning=FALSE, fig.width=8, fig.height=5, fig.show = "hold", out.width = "290px"}
dsub <- d %>% select("Country", "Year", "Publisher", "Cost") %>%
              filter(Publisher %in% selected.publishers) %>%
	      group_by(Country, Year, Publisher) %>%
	      summarize(Cost = sum(Cost, na.rm = TRUE))


# ggplot(dsub, aes(x = Year, y = Cost, fill = Publisher)) + facet_grid(Country ~ .) + geom_bar(stat = "identity", position = "dodge") + scale_y_log10()
for (publisher in c("Elsevier", "Springer", "Wiley / Blackwell")) {
  p <- ggplot(subset(dsub, Publisher == publisher), aes(x = Year, y = Cost, color = Country)) + geom_line() + geom_point() + scale_y_log10() + labs(title = publisher, y = "Cost (EUR)")
  print(p)
}
```

`r figs("costs1", "Cost timelines for selected publishers.")`




```{r countries.with.data, cache=FALSE, echo=FALSE, message=FALSE, warning=FALSE, fig.width=10, fig.height=10, fig.show = "hold", out.width = "400px"}
par(mar=c(3, 16, 3, 2));
n <- rev(sort(colSums(table(d$Country, d$Publisher) > 0)));
tab <- rev(n[n>=2])
# Pick publishers with multi-country data
selected.publishers <- names(tab)
barplot(tab, las = 1, horiz = T, main = "Number of countries with data")
```

`r figs("countries.with.data", "Number of countries with data for at least one year. Publishers with more than one country are shown.")`







